Which Scenario Should We Design For? Insights from House Elevation on the Multiple PDF Problem

### 2021-12-10 papers/agu21/case-study.tex
As printed for AGU!

\begin{block}{case study: norfolk, va}
    We model a \emph{hypothetical} house in Norfolk, VA, where sea level rise drives nonstationary future flood hazard, following the approach of \cite{zarekarizi_suboptimal:2020}.
    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{historic_surge.pdf}
        \caption{
            Time series of annual maximum of storm surge (after subtracting mean sea level) at Sewells Point, VA (NAVD datum).
        }
        \label{fig:observations}
    \end{figure}
    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{hazard-tradeoffs.pdf}
        \caption{
            Neglecting hydrodynamics, we add sea level rise (a; see ref.~\cite{ruckert_coastal:2019}) to a Bayesian GEV model of storm surge (b) to get flood hazard.
            The convolution of hazard and fragility (c; see ref.~\cite{zarekarizi_suboptimal:2020}) yields an assessment of damages and tradeoffs (\cref{fig:tradeoffs}).
        }
    \end{figure}
\end{block}

>>>>

\begin{block}{case study: norfolk, va}
    We model a \emph{hypothetical} house in Norfolk, VA, where sea level rise drives nonstationary future flood hazard, following the approach of ref.~\cite{zarekarizi_suboptimal:2020}.
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=\textwidth]{historic_surge.pdf}
            \caption{
                Time series of annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA.
                Purple (orange) arrows denote notable tropical cyclones (nor'easters).
            }
            \label{fig:observations}
        \end{figure}
    \end{framed}
    \vspace*{-1em}
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=\textwidth]{hazard-tradeoffs.pdf}
            \caption{
                Neglecting hydrodynamics, we add sea level rise (a) to a Bayesian GEV model of storm surge (b) to get flood hazard.
                The convolution of hazard and fragility (c) yields an assessment of damages and tradeoffs (\cref{fig:tradeoffs}).
            }
        \end{figure}
    \end{framed}
\end{block}

### 2021-12-10 papers/agu21/combining.tex
As printed for AGU!

\begin{block}{synthesizing pdfs for decision relevance}
    If there are only a few PDFs, then qualitative comparisions (\eg, \cref{fig:tradeoffs}) may be sufficient.
    With \emph{many} PDFs (\eg, \cref{fig:boxplots}), further synthesis is needed.

    We consider the \emph{common case} of assessing decisions using simulations (SOWs) from each of $K$ PDFs.
    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{bayes-rdm.pdf}
        \caption{
            Sketch of the ``Prior Model Averaging'' framework.
        }\label{fig:flowchart}
    \end{figure}
\end{block}

>>>>

\begin{block}{synthesizing pdfs for decision relevance}
    While comparing results across many PDFs is a valuable exploratory exercise (see, \eg, \cref{fig:tradeoffs}), providing end users (\eg, households practicing engineers, and planning departments of local governments) with many PDFs may lead to confusion, inconsistency, and potentially liability (see refs.~\cite{schneider_dangerous:2001,schneider_scenarios:2002}).
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=0.8\textwidth]{bayes-rdm.pdf}
            \caption{
                ``Prior Model Averaging'' framework.
            }\label{fig:flowchart}
        \end{figure}
    \end{framed}
    Our approach to synthesizing PDFs is designed for the common case of assessing decisions using simulations (SOWs) from each of $K$ PDFs.
    A system model $(f)$ quantifies the performance $(u)$ of a decision ($x$) under a single SOW ($s$).
    To synthesize, we weight PDFs using an approach based on Bayesian Model Averaging \cite{wong_surge:2018,Yao:2018bu}.
\end{block}

### 2021-12-10 papers/agu21/doesnt-fit-all.tex
As printed for AGU!

\begin{block}{one size doesn't fit all!}
    \begin{itemize}
        \item Tailoring guidance to specific building characteristics can improve outcomes
        \item Elevating to ``\gls{bfe} plus a foot'' is not always optimal \cite{xian_elevation:2017,zarekarizi_suboptimal:2020}
        \item Both over- and under-building can be costly \cite{ansar_bigisfragile:2017,DossGollin:2019}
    \end{itemize}
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=0.475\textwidth]{5.5.pdf}%
            \includegraphics[width=0.475\textwidth]{7.5.pdf}
            \caption{
                Tradeoffs between construction cost and expected NPV lifetime costs for houses initially situated (L) \SI{2.5}{ft} and (R) \SI{0.5}{ft} below the \gls{bfe} under \gls{rcp} 6.0 with fast BRICK dynamics.
                Note $y$-axes are not equal.
            }
        \end{figure}
    \end{framed}
\end{block}

>>>>

\begin{block}{one size does not fit all!}
    \begin{itemize}
        \item Tailoring guidance to specific building characteristics can improve outcomes
        \item Elevating to ``\acrshort{bfe} plus a foot'' is not always optimal \cite{xian_elevation:2017,zarekarizi_suboptimal:2020}
        \item Both over- and under-building can be costly \cite{ansar_bigisfragile:2017,DossGollin:2019}
    \end{itemize}
    \begin{framed}
        \begin{figure}
            \centering
            \begin{subfigure}{0.48\textwidth}
                \subcaption{\SI{2.5}{ft} below BFE}
                \includegraphics[width=\textwidth]{5.5.pdf}
            \end{subfigure}
            \hfill
            \begin{subfigure}{0.48\textwidth}
                \subcaption{\SI{0.5}{ft} below BFE}
                \includegraphics[width=\textwidth]{7.5.pdf}
            \end{subfigure}
            \caption{
                Tradeoffs between construction cost and expected NPV lifetime costs for houses initially situated (a) \SI{2.5}{ft} and (b) \SI{0.5}{ft} below the \gls{bfe} (under \gls{rcp} 6.0 with fast BRICK \cite{wong_brick0.2:2017} dynamics).
                Note unequal $y$-axis scales.
            }
        \end{figure}
    \end{framed}
\end{block}

### 2021-12-10 papers/agu21/house-elevation.tex
As printed for AGU!

\begin{block}{house elevation}
    Households elevate their homes to manage flood risks, but regulations and guidance are silent on key questions \cite{zarekarizi_suboptimal:2020,xian_elevation:2017}.
    \begin{framed}
        \begin{description}
            \item[Q1] How to adapt guidance to building characteristics or household preferences?
            \item[Q2] How does nonstationary hazard change guidance?
            \item[Q3] Which model of nonstationary hazard should be used?
        \end{description}
    \end{framed}
    \begin{figure}
        \centering
        \subcaptionbox{Selinsgrove, PA}{\includegraphics[height=3.35in]{img_selinsgrove.jpeg}}%
        \hfill
        \subcaptionbox{Houston, TX}{\includegraphics[height=3.35in]{img_braeswood.png}}\\
        \subcaptionbox{Plaquemines Parish, LA}{\includegraphics[height=3.3in]{epstein-2020-la.jpeg}}%
        \hfill
        \subcaptionbox{Dare County, NC}{\includegraphics[height=3.3in]{nichols-ocracoke.jpeg}}%
        \caption{
            \faIcon{camera}: (a) JDG, (b) Google Maps, (c) Mitch Epstein /  New York Times (d) Rob Nichols.
        }
        \label{fig:images}
    \end{figure}
\end{block}

>>>>

\begin{block}{house elevation}
    Households elevate their homes to manage flood risks, but regulations and guidance are silent on key questions \cite{zarekarizi_suboptimal:2020,xian_elevation:2017}.
    \vspace*{-1em}
    \begin{description}
        \item[Q1] How to adapt guidance to building characteristics or household preferences?
        \item[Q2] How does nonstationary hazard change guidance?
        \item[Q3] Which model of nonstationary hazard to use?
    \end{description}
    \begin{framed}
        \begin{figure}
            \centering
            \subcaptionbox{Selinsgrove, PA}{\includegraphics[height=3.35in]{img_selinsgrove.jpeg}}%
            \hfill%
            \subcaptionbox{Houston, TX}{\includegraphics[height=3.35in]{img_braeswood.png}}\\
            \subcaptionbox{Plaquemines Parish, LA}{\includegraphics[height=3.3in]{epstein-2020-la.jpeg}}%
            \hfill%
            \subcaptionbox{Dare County, NC}{\includegraphics[height=3.3in]{nichols-ocracoke.jpeg}}%
            \caption{
                \faIcon{camera}: (a) JDG, (b) Google Maps, (c) Mitch Epstein /  New York Times (d) Rob Nichols / PSU.
            }
            \label{fig:images}
        \end{figure}
    \end{framed}
    These challenges also affect other adaptation plans and engineering designs (\cref{fig:stormwater-bridge}).
\end{block}

### 2021-12-10 papers/agu21/multiple-pdf.tex
As printed for AGU!

\begin{block}{the multiple pdf problem}
    Uncertainties from multiple model structures (\eg, ice sheet dynamics) and/or scenarios (\eg, emissions pathways) create multiple estimates of time-varing hazard, \ie the \textbf{\itshape{Multiple PDF Problem}}.
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=\textwidth]{msl_boxplots.pdf}
            \caption{
                Sea level rise projections use the BRICK model \cite{wong_brick0.2:2017}.
                We consider \emph{eight time-dependent PDFs of sea level rise}: four \gls{rcp} scenarios $\times$ two parameterizations of ice sheet processes.
            }\label{fig:boxplots}
        \end{figure}
    \end{framed}
    Considering other models \cite{kopp_evolving:2017,deconto_antarctica:2016,ruckert_coastal:2019} would \emph{compound} this problem.
    \textbf{Which PDF should long-lasting infrastructure like housing be designed for?}
\end{block}

>>>>

\begin{block}{the multiple pdf problem}
    Uncertainties from multiple model structures (\eg, ice sheet dynamics) and/or scenarios (\eg, emissions pathways) create multiple estimates of time-varing hazard, \ie, the \textbf{\itshape{Multiple PDF Problem}}.
    Including additional models \cite{kopp_evolving:2017,deconto_antarctica:2016,ruckert_coastal:2019} \emph{would compound this challenge}.
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=\textwidth]{msl_boxplots.pdf}
            \caption{
                Sea level rise projections use the BRICK model \cite{wong_brick0.2:2017,ruckert_coastal:2019}.
                We consider \emph{eight time-dependent PDFs of sea level rise}: four \gls{rcp} scenarios $\times$ two parameterizations of ice sheet processes.
            }\label{fig:boxplots}
        \end{figure}
    \end{framed}
\end{block}

### 2021-12-10 papers/agu21/not-equally-plausible.tex
As printed for AGU!

\begin{block}{we're always }
    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{lsl_priors.pdf}
        \caption{
            If we weight all scenarios equally, then our projection of MSL in 2100 follows the sampling distribution.
            This may or may not accurately reflect available information.
        }
    \end{figure}
    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{inference_weights.pdf}
        \caption{
            The prior model averaging approach (\cref{fig:flowchart}) re-weights PDFs based on their distance, and that of other PDFs available, from the (ideally coproduced!) prior.
        }
    \end{figure}
\end{block}

>>>>

\begin{block}{key message}
    \vspace*{-1em}
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=0.48\textwidth]{lsl_priors.pdf}~
            \includegraphics[width=0.48\textwidth]{tradeoffs_height_totalcost_byprior.pdf}~
            \caption{
                The prior model averaging approach (\cref{fig:flowchart}), uses a probability distribution representing subjective belief to average insights across each PDF available.
                Alternative priors can provide robustness checks.
            }
        \end{figure}
    \end{framed}
    \begin{itemize}
        \item Natural extensions to Bayesian updating
        \item Belief $p(s)$ can't be ``right''  \cite{gelman_workflow:2020,gelman_philosophy:2013} or ``neutral'' \cite{quinn_exploratory:2020}
        \item Instead, we can ask \textbf{what assumptions are different priors consistent with?}
    \end{itemize}
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=\textwidth]{inference_weights.pdf}
            \caption{
                The weights assigned to each PDF (from \cref{fig:boxplots}) under each prior considered.
            }
        \end{figure}
    \end{framed}
    ``Pessimistic'' (``Optimistic'') prior heavily weights \gls{rcp} 8.5 (2.6), which is unlikely given current policies \cite{hausfather_scenarios:2020}.
\end{block}

### 2021-12-10 papers/agu21/other-applications.tex
As printed for AGU!

\begin{block}{looking ahead}
    This didactic example illustrates a need for better synthesis and communication of deep uncertainties for decision making.
    \textbf{\scshape let's collaborate on:}
    \begin{itemize}
        \item More complex models to capture more relevant metrics
        \item Better priors over nonstationary hazard
        \item Interacting, sequential decisions
        \item Inclusive assessment of ethical questions around scenario weighting
    \end{itemize}
    \begin{framed}
        \begin{figure}
            \centering
            \subcaptionbox{Stormwater}{\includegraphics[height=3.5in]{culvert.jpeg}}%
            \hfill
            \subcaptionbox{Bridge Clearance}{\includegraphics[height=3.5in]{bridge_clearance.jpg}}
            \caption{
                House elevation is just one of many design problems where nonstationary hazards, subject to scenario and model structure uncertainties, drive outcomes.
                \faIcon{camera}: (a) JDG (b) D. Howard / Wikipedia.
            }
            \label{fig:images}
        \end{figure}
    \end{framed}
\end{block}

>>>>

\begin{block}{looking ahead}
    This didactic example illustrates a need for better synthesis and communication of deep uncertainties for decision making.
    \textbf{\scshape let's collaborate on:}
    \begin{itemize}
        \item More complex models to capture more relevant metrics
        \item Better models of nonstationary hazards
        \item Interacting, sequential decisions
    \end{itemize}
    \begin{framed}
        \begin{figure}
            \centering
            \subcaptionbox{Stormwater}{\includegraphics[height=3.5in]{culvert.jpeg}}%
            \hfill
            \subcaptionbox{Bridge Clearance}{\includegraphics[height=3.5in]{bridge_clearance.jpg}}
            \caption{
                House elevation is just one of many design problems where nonstationary hazards, subject to scenario nd model structure uncertainties, drive outcomes.
                \faIcon{camera}: (a) JDG (b) D. Howard / Wikipedia.
            }
            \label{fig:stormwater-bridge}
        \end{figure}
    \end{framed}
\end{block}

### 2021-12-10 papers/agu21/scenario-matters.tex
As printed for AGU!

\begin{block}{it matters which pdf you choose!}
    Any optimal strategy or Pareto frontier is conditional upon an (explicit or implicit) model of future outcomes (see ref.~\cite{gelman_philosophy:2013} for a more philosophical discussion).
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=0.48\textwidth]{tradeoffs_height_totalcost_byscenario.pdf}%
            \includegraphics[width=0.48\textwidth]{tradeoffs_height_iqr_byscenario.pdf}
            \caption{
                Tradeoffs between (L) construction cost and expected lifetime costs or (R) construction cost and the uncertainty of future lifetime costs depend on the model selected.
            }
            \label{fig:tradeoffs}
        \end{figure}
    \end{framed}
    Weights assigned to PDFs or simulations, \textbf{including implicit uniform weights}, should be communicated transparently to facilitate critique and improvement
\end{block}

>>>>

\begin{block}{it matters which pdf you choose!}
    Any optimal strategy or Pareto frontier is conditional upon an (explicit or implicit) model of future outcomes.
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=0.48\textwidth]{tradeoffs_height_totalcost_byscenario.pdf}%
            \includegraphics[width=0.48\textwidth]{tradeoffs_height_iqr_byscenario.pdf}
            \caption{
                Tradeoffs between (L) construction cost and expected lifetime costs or (R) construction cost and the uncertainty of future lifetime costs depend on the PDF selected.
            }
            \label{fig:tradeoffs}
        \end{figure}
    \end{framed}
    Weights assigned to PDFs or simulations, \textbf{including implicit uniform weights}, should be communicated transparently to facilitate critique and improvement
\end{block}
